---
title: "WT:WP1:Paper 1"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    code-link: true
    embed-resources: true
    page-layout: full
    grid:
      margin-width: 100px
      body-width: 1000px
author: Daniel Väisänen
editor:
  markdown: 
    wrap: sentence
toc: true
execute: 
  warning: false
  echo: true
  freeze: auto
highlight-style: custom.theme
theme:
  light: github
  dark: darkly
---
# Project description

Paper 1 in work project 1 of the work together series

#-------------------------
## Libraries

```{r}
xfun::pkg_attach(
  "tidyverse", 
  "dtrackr", 
  "glue", 
  "DT", 
  "flextable"
  )

```
#-------------------------
## Base theme

```{r}
theme_set(
  theme_minimal(
    base_size = 16,
    base_family = "Roboto"
  ) +
    theme(
      legend.position = "top",
      panel.grid.major.y = element_line(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.x = element_line(),
      panel.grid.minor.x = element_blank(),
      axis.title.y = element_text(margin = margin(r = 10), colour = "grey30"),
      axis.title.x = element_text(margin = margin(t = 10), colour = "grey30")
    )
)
```

#-------------------------
## Data

```{r}
hpa <- read_csv(here::here("..", "..","..", "HPI (DNR XXXXXX)", "Data HPI registeruttag 2023", "HPI", "HPI clean", "EEB_hpb_clean_2023-10-16.csv")) %>% 
    mutate(SymptomBackNeck = if_else(SymptomBackNeck == 7 | SymptomBackNeck == 8, NA, SymptomBackNeck)) 

scb <- read_csv(here::here("..", "..","..", "HPI (DNR XXXXXX)", "Data HPI registeruttag 2023", "SCB", "SCB clean", "Grunduppgifter_2023-09-24.csv")) %>% mutate(Kommun=str_replace(Kommun, "^0+", ""))# remove leading zeros

# join SCB and HPI data
df <- hpa %>% inner_join(scb, by = join_by(LopNr, Performed, Year))  %>% 
    mutate(SSYK1_scb_hpi_combined = coalesce(SSYK1, Ssyk1_hpi_combined))

```


# Selection
different n depending on different selection criterias
```{r}

# Assuming 'hpa' is your original data frame
# Calculate base counts
base_counts <- hpa %>% 
  arrange(LopNr, Performed) %>% 
  mutate(hpa_number = row_number(), .by = LopNr) %>% 
  count(hpa_number) %>% 
  slice_head(n = 10) %>%
  rename(base_count = n)

# Calculate counts for different conditions
hpa_counts <- hpa %>% 
  arrange(LopNr, Performed) %>% 
  mutate(hpa_number = row_number(), .by = LopNr)

# Function to filter and count based on minimum number of non-NA values
count_non_na <- function(data, phys_min, quest_min) {
  data %>% 
    filter(
      rowSums(!is.na(select(., starts_with("EkB_rel_VO2"), starts_with("Astrand_rel_VO2"), 
                             starts_with("HeightCM"), starts_with("WeightKG"), 
                             starts_with("BloodPressureSystolic"), starts_with("BloodPressureDiastolic")))) >= phys_min,
      rowSums(!is.na(select(., starts_with("ExerciseAnswer"), starts_with("TobaccoSmoking"), 
                             starts_with("Diet"), starts_with("Health"), 
                             starts_with("StressOverall"), starts_with("SymptomBackNeck")))) >= quest_min,
      Year > 1994
    ) %>% 
    count(hpa_number) %>% 
    slice_head(n = 10) %>%
    # Create a column with a dynamic name based on phys_min and quest_min
    mutate("{paste0('n', phys_min, '_', quest_min)}" := n) %>%
    select(-c(hpa_number, n))
}

# Apply the function with different minimum values
counts_1_1 <- count_non_na(hpa_counts, 1, 1) 
counts_1_2 <- count_non_na(hpa_counts, 1, 2)
counts_2_1 <- count_non_na(hpa_counts, 2, 1)
counts_2_2 <- count_non_na(hpa_counts, 2, 2)
counts_3_2 <- count_non_na(hpa_counts, 3, 2)
counts_2_3 <- count_non_na(hpa_counts, 2, 3)
counts_3_3 <- count_non_na(hpa_counts, 3, 3)

# Join all counts together
sums <- bind_cols(base_counts, counts_1_1, counts_1_2, counts_2_1, counts_2_2, counts_3_2, counts_2_3, counts_3_3) 
  

sums %>%
  mutate(
    diff_1_1 = base_count - n1_1,
    ande_1_1 = (diff_1_1 / base_count) * 100,
    diff_1_2 = base_count - n1_2,
    ande_1_2 = (diff_1_2 / base_count) * 100,
    diff_2_1 = base_count - n2_1,
    ande_2_1 = (diff_2_1 / base_count) * 100,
    diff_2_2 = base_count - n2_2,
    ande_2_2 = (diff_2_2 / base_count) * 100,
    diff_3_2 = base_count - n3_2,
    ande_3_2 = (diff_3_2 / base_count) * 100,
    diff_2_3 = base_count - n2_3,
    ande_2_3 = (diff_2_3 / base_count) * 100,
    diff_3_3 = base_count - n3_3,
    ande_3_3 = (diff_3_3 / base_count) * 100
  ) %>% select(-starts_with("n")) %>% 
  mutate(across(everything(), round(1))) %>% 
  flextable::flextable()




```



# Variable creations
* Uppgift saknas
1 Förgymnasial utbildning kortare än 9 år
2 Förgymnasial utbildning 9 år
3 Gymnasial utbildning högst 2-årig
4 Gymnasial utbildning 3 år
5 Eftergymnasial utbildning kortare än 3 år
6 Eftergymnasial utbildning 3 år eller längre (exkl. forskarutbildning)
7 Forskarutbildning 


High-Skilled Group:  
Managers and Administrators
Professionals (e.g., healthcare, education, IT)
Technicians and Associate Professionals
This group encompasses roles that typically require advanced education, specialized skills, and a high level of expertise.

Skilled Manual (Blue Collar) Group:  
Skilled Agricultural, Forestry, and Fishery Workers
Craft and Related Trades Workers
Plant and Machine Operators, and Assemblers
Jobs in this group generally require specific vocational skills or training, often involving manual labor or technical expertise.
Clerical and Service Workers Group:

Clerical Support Workers:   
Service and Sales Workers
This group covers roles primarily in the service sector, including both office-based clerical jobs and customer-facing service roles. These occupations may require certain levels of training and skills, but not to the extent of the high-skilled group.
Elementary Occupations Group:

Elementary Occupations:    
This group includes occupations that usually perform simple and routine tasks, which may require the use of hand-held tools and physical effort. These roles often do not require formal educational credentials or specialized training.
```{r}
# data from tillväxtverket, another possibility could be skr
metropolitan_municipalities <- c("114", "123", "126", "127", "128", "136", "138", "139", "160", 
                                 "162", "163", "180", "181", "182", "183", "184", "186", "191", 
                                 "1230", "1231", "1262", "1280", "1281", "1402", "1480", "1481")

dense_municipalities <- c("115", "117", "120", "140", "187", "305", "330", "380", "381", "461", 
                          "480", "481", "483", "484", "486", "561", "562", "580", "581", "583", 
                          "584", "642", "643", "680", "682", "683", "686", "687", "760", "765", 
                          "780", "781", "880", "881", "882", "883", "884", "1060", "1080", "1082", 
                          "1233", "1261", "1263", "1272", "1275", "1282", "1277", "1282", "1283", 
                          "1285", "1286", "1287", "1290", "1292", "1380", 
                          "1382", "1401", "1407", "1440", "1441", "1472", "1482", "1484", "1485", 
                          "1486", "1487", "1488", "1489", "1490", "1492", "1493", "1494", "1495", 
                          "1496", "1497", "1498", "1499", "1715", "1761", "1780", "1781", "1782", 
                          "1784", "1785", "1862", "1880", "1881", "1883", "1884", "1907", "1960", 
                          "1961", "1980", "1981", "1982", "1983", "1984", "2062", "2080", "2081", 
                          "2085", "2104", "2180", "2181", "2262", "2280", "2281", "2380", "2523", "2580", "2581", 
                          "2582", "2583", "2584")

rural_municipalities <- c("125", "128", "188", "192", "319", "331", "360", "382", "428", "482", "488", "509", 
                          "512", "513", "560", "563", "580", "582", "586", "604", "617", "662", "665", 
                          "684", "685", "761", "763", "764", "767", "821", "834", "840", "860", "861", "862", "885", 
                          "980", "1081", "1083", "1214", "1256", "1257", "1260", "1264", "1265", 
                          "1266", "1267", "1270", "1273", "1276", "1278", "1284", "1291", "1293", "1315", 
                          "1381", "1383", "1384", "1415", "1419", "1421", "1427", "1430", "1435", 
                          "1438", "1439", "1442", "1443", "1444", "1445", "1446", "1447", "1452", 
                          "1460", "1461", "1462", "1463", "1465", "1466", "1491", "1470", "1471", "1473", 
                          "1730", "1737", "1760", "1762", "1763", "1764", "1765", "1766", "1783", "1814", 
                          "1860", "1861", "1863", "1864", "1882", "1885", "1904", "1962", "2021", 
                          "2023", "2026", "2029", "2031", 
                          "2034", "2039", "2061", "2082", "2083", "2084", "2101", "2121", "2132", 
                          "2161", "2182", "2183", "2184", "2260", "2282", "2283", "2284", "2303", 
                          "2305", "2309", "2313", "2321", "2326", "2361", "2401", "2403", "2404", 
                          "2409", "2417", "2418", "2421", "2422", "2425", "2460", "2462", "2463", 
                          "2480", "2481", "2482", "2505", "2506", "2510", "2513", "2514", "2518", 
                          "2521", "2560")






df1 <- df %>% 
   mutate(Kommun=str_replace(Kommun, "^0+", "")) %>% 
  # Convert Year to date and create 5-year periods
  mutate(Period = case_when(
    Year >= 1995 & Year < 2000 ~ "1995-1999",
    Year >= 2000 & Year < 2005 ~ "2000-2004",
    Year >= 2005 & Year < 2010 ~ "2005-2009",
    Year >= 2010 & Year < 2015 ~ "2010-2014",
    Year >= 2015 & Year < 2020 ~ "2015-2019",
    Year >= 2020 & Year <= 2023 ~ "2020-2023",
    TRUE ~ NA_character_  # For years outside the specified range
  )) %>%
  # Create age groups
  mutate(AgeGroup = case_when(
    Age >= 18 & Age <= 35 ~ "18-35",
    Age > 35 & Age <= 50 ~ "36-50",
    Age > 50 & Age <= 65 ~ "51-65",
    TRUE ~ ">65"
  )) %>%
  # Categorize Civil status
    mutate(MarriedOrPartner = case_when(
    Civil %in% c("G", "RP") ~ "Partner",  # Married or Registrerad Partner
    Civil %in% c("OG", "S", "Ä", "SP", "EP") ~ "NonPartnered",  # Other categories (you can specify these)
    TRUE ~ NA_character_  # Assign NA to cases that do not match any condition
  )) %>% 
  # Categorize Birth country
  mutate(BirthPlace = case_when(
    FodelseLandEU28 == "Sverige" ~ "Sweden",
    FodelseLandEU28 %in% c("EU utom Norden", "Europa utom EU och Norden", "Norden utom Sverige") ~ "In Europe",
    TRUE ~ "Outside Europe"
  )) %>%
  # Categorize Kommun based on hypothetical population size categories
  mutate(KommunSize = case_when(
    Kommun %in% metropolitan_municipalities ~ "metropolitan_municipalities",
    Kommun %in% dense_municipalities ~ "dense_municipalities",
    Kommun %in% rural_municipalities ~ "rural_municipalities",
    TRUE ~ "Other"
  )) %>%

  # Categorize education level
  mutate(EducationLevel = case_when(
    SUN2000Niva_Old %in% 1:2 ~ "Primary",
    SUN2000Niva_Old %in% 3:4 ~ "Secondary",
    SUN2000Niva_Old %in% 5 ~ "Tertiary <2 years",
    SUN2000Niva_Old %in% 6:7 ~ "Higher Education",
    TRUE ~ "Unknown"
  )) %>% 
    # Divide income into quartiles
  mutate(IncomeQuartile = ntile(CSFVI, 4), 
         .by = Period) %>%
   mutate(IncomeLevel = case_when(
    CSFVI >= 2 * median(CSFVI) ~ "≥200%",
    CSFVI >= 1.2 * median(CSFVI) & CSFVI < 2 * median(CSFVI) ~ "120–199%",
    CSFVI >= 0.8 * median(CSFVI) & CSFVI < 1.2 * median(CSFVI) ~ "80–119%",
    CSFVI >= 0.6 * median(CSFVI) & CSFVI < 0.8 * median(CSFVI) ~ "60–79%",
    TRUE ~ "<60%"
  ), .by = Period) %>% 
  # combine occupation
    mutate(
  SSYK1_scb_hpi_combined = coalesce(SSYK1, Ssyk1_hpi_combined),
  # Categorize occupation
    SSYK_WB = case_when(SSYK1_scb_hpi_combined <= 4 ~ "White-Collar",
                             SSYK1_scb_hpi_combined > 4 ~ "Blue-Collar"),
SSYK_Category = case_when(
    SSYK1_scb_hpi_combined %in% 1:3 ~ "High-Skilled",                        # High-skilled group
    SSYK1_scb_hpi_combined %in% 4:5 ~ "Clerical and Service Workers",        # Clerical and service workers group
    SSYK1_scb_hpi_combined %in% 6:8 ~ "Skilled Manual",        # Skilled manual or blue-collar group
    SSYK1_scb_hpi_combined == 9 ~ "Elementary Occupations"                   # Elementary occupations group
)
  )

```

# Flowchart and data in/ex-clusion

```{r}

# global options for dtrackr

old = options(
 #dtrackr.strata_glue="{.group}:{.value}",
  #dtrackr.strata_sep="; ",
  dtrackr.strata_glue="{tolower(.value)}",
  dtrackr.strata_sep=", ",
  dtrackr.default_message = "n={formatC(.count, big.mark = ',',format = 'fg')}"
  #dtrackr.default_headline = "subgroup: {.strata}"
)
# reset options 
#options(old)

df2 <-
df1 |> 
  track(.headline = "Flow chart     ",
        .messages="Data SCB variables in HPI data"
  ) %>%
  status(.messages = "n={formatC(.count, big.mark = ',')}") %>%  
  filter(rowSums(!is.na(select(., EkB_rel_VO2, Astrand_rel_VO2, HeightCM, WeightKG, BloodPressureSystolic, BloodPressureDiastolic))) >= 2,
         rowSums(!is.na(select(., ExerciseAnswer, TobaccoSmoking, Diet, Health, StressOverall, SymptomBackNeck))) >= 2, 
         .messages = "Excluded {formatC(.excluded, big.mark = ',',format = 'fg')}", .headline = "questionnaire 2 physilogical 2" ) %>% 
   status(.messages = "n={formatC(.count, big.mark = ',')}") %>% 
  exclude_all(
    Year < 1995 ~ "{formatC(.excluded, big.mark = ',',format = 'fg')} Year < 1995",
    CSFVI == 0 ~ "{formatC(.excluded, big.mark = ',',format = 'fg')} with income 0kr",
   # Age > 65 ~  "{formatC(.excluded, big.mark = ',',format = 'fg')} with age > 65" ,
    is.na(SSYK1_scb_hpi_combined) ~  "{formatC(.excluded, big.mark = ',',format = 'fg')} with NA in SSYK" ,
    is.na(Civil) ~  "{formatC(.excluded, big.mark = ',',format = 'fg')} with NA in Civil",
    FodelseLandEU28=="Okänt" ~  "{formatC(.excluded, big.mark = ',',format = 'fg')} with 'Okänt'in fodelselandEU28",
   # SUN2000Niva_Old=="*" ~  "{formatC(.excluded, big.mark = ',',format = 'fg')} with NA in education (*)"
  ) %>% 
  status(.messages = "n={formatC(.count, big.mark = ',')}") %>% 
      arrange(LopNr, Performed) %>% 
  mutate(hpa_number = row_number(), 
         test_count = n(), .by = LopNr) %>% 
  filter(hpa_number == 1,
         .messages = "Excluded {formatC(.excluded, big.mark = ',',format = 'fg')}", .headline = "Repeated tests") %>% 
 # status() %>% 
  group_by(test_count>1,) %>% 
  comment(" has {formatC(.count, big.mark = ',')} items") %>%
  ungroup(.headline = "Total study population\nn={formatC(.count, big.mark = ',')}",
          .messages = "") 



df1 %>% flowchart()


#df1 |> 
 # flowchart(filename="testtype",
  #          formats = c("png","svg", "pdf"),
   #                      maxWidth=20, 
    #                    maxHeight = 20)



```

#-------------------------
# Descriptive tables

## function for creating tables

```{r Table functions}
# test statistics
# continuous multi group: anova/Kruskal  (parametric/non-parametric)
# continuous one or two group: t-test/Man-Whitney U (parametric/non-parametric)
# categorical two or more groups: Pearson's chi-squared or Fisher’s exact test
# if any of the expected frequencies are < 5 in the contingency table

# https://www.saem.org/about-saem/academies-interest-groups-affiliates2/cdem/for-students/cdem-voice/educational-research-column/educational-research-column-choosing-wisely-chi-square-vs.-fisher-s-exact

# https://statsandr.com/blog/chi-square-test-of-independence-in-r/#chi-square-test-of-independence-in-r

###################
# FACTOR FUNCTION #
###################
# df <- df
# var <- "Age"
# var <- "Sex"
# group <- "Period"
# val <- c(Age", "Sex")

# d <- factor.fun(df, Contraception, ModifiedM4)
# d <- factor.fun(df, BV_Diagnosis_v3, ModifiedM4) 
# d <- factor.fun(df, `Below LLD Estradiol`, Contraception) 
factor.fun <- function(df, var, group){
  var <- enquo(var)
  group <- enquo(group)
  
  contingency.table <- df %>%
    select(!!var, !!group) %>%
    table() 
    
  # Pearson’s Chi-squared test:
  test <- chisq.test(contingency.table)
  t <- "2"
  
  # Fisher’s exact test:
  if(min(test$expected)<5)
    {test <- fisher.test(contingency.table)
    t <- "4"}
    
  p.val <- test[["p.value"]] #%>% formatC(., format = "e", digits = 1) %>% paste0("$$ ",.,"^2 $$")
  if(p.val == ""){p.val <- NA}
    else if(p.val < 0.01){p.val <- "<0.01"}
      else{p.val <- sprintf(p.val, fmt = '%#.2f')}
  na <- sum(is.na(pull(df, !!var)))
  tab <- tibble("Variables"=as_label(var),"na"=as.numeric(na), "p-value"=p.val) #as.numeric(p.val)
  
  # mean, min and max:
  sum_tab <- df %>%
    select(!!var, !!group) %>%
    group_by(!!group, !!var) %>%
    summarize(n = n(), .groups = "drop_last") %>%
    nest() %>%
    ungroup() %>%
    mutate(sum = map(data, ~sum(.$n))) %>%
    unnest(c(data, sum)) %>%
    mutate(percent = round(100*(n/sum))) %>%
    mutate( val = paste0(format(.$n, big.mark = ",")," (",.$percent,"%)")) %>% # !!as_label(var) :
    select(-sum, -n, -percent) %>%
    pivot_wider(., names_from = !!group, values_from = val ) %>%
    dplyr::rename(Variables = as_label(var)) %>%
    arrange(Variables) %>%
    mutate_at(vars(Variables), ~replace(., is.na(.), "Not available")) %>%
    mutate_at(vars(Variables), ~paste0("  - ", .)) %>%
    mutate_if(is.character, ~replace(., is.na(.), 0))

   t <- bind_rows(tab, sum_tab) %>% mutate("t" = t)
  
  return(t)
  #return(sum_tab)
  #return(list(tab, na, p.val, sum_tab))
}


factor.fun(df=df2 %>% select(Astrand_rel_VO2, Age, Sex, Period), 
           var = c("Sex"),
           group = Period)

#######################
# CONTINUOUS FUNCTION #
#######################
# d_ <- continuous.fun(df, Age, Period)
continuous.fun <- function(df, var, group, stat_type = "mean_ci"){
  var <- enquo(var)
  group <- enquo(group)
  name_g <- as_label(group) 
  name_v <- as_label(var) 
  
  # round decimals
  round.fun <- function(x){
    df <- x %>%
    mutate_if(
    is.numeric,
    ~ ifelse(.x >= 1|.x==0, round(.x, digits = 1),
            ifelse(.x<= 0.9 & .x >= 0.01, round(.x, digits = 2),
                                          formatC(.x, format = "e", digits = 1))))
  }
  
    # Helper function for numeric summary (mean and 95% CI)
  mean_confidence.fun <- function(m, se) {
    m <- m
    se <- se
    ci_lower <- round(m - 1.96 * se, 2)
    ci_upper <- round(m + 1.96 * se, 2)
    return(paste0(m, " (", ci_lower, "-", ci_upper, ")"))
  }
  # make factors
  if( is.null(levels(group)) ){df <- mutate( df, !!(name_g) := factor((!!group)) ) }

  # p-value
  exp1 <- expr(!!ensym(var) ~ !!ensym(group))
  if( length(unique(pull(df, !!group))) > 2 ){
    test <- kruskal.test(formula = eval(exp1), data = df) 
    p.val <- test[["p.value"]]
    if(p.val < 0.01){p.val <- "<0.01"}else{p.val <- sprintf(p.val, fmt = '%#.2f')}
    t <- "3"
  }else{
    lvl <- df %>% select(!!var, !!group) %>% na.omit() %>% pull(., (!!group)) %>% unique() %>% length()
    if(lvl == 2){
    test <- wilcox.test(formula = eval(exp1), data = df)
    p.val <- test[["p.value"]]
    if(p.val < 0.01){p.val <- "<0.01"}else{p.val <- sprintf(p.val, fmt = '%#.2f')}
    t <- "1"}
    if(lvl < 2) {
      p.val <- NA
      t <- "5"
    }
  }
  l <- unique(pull(df, !!group))
  #l <- c("DMPA", "no HC")
  
  # mean, min max or similar
  #if(stat_type == "median")else{stat_type == "mean"}
  sum_tab <- df %>%
    select(!!var, !!group) %>%
    mutate(na = sum(is.na(!!var))) %>%
    group_by(!!group) %>%
    summarize(median = median((!!var),na.rm = T),
              mean = mean((!!var),na.rm = T),
              se = sd((!!var), na.rm = TRUE) / sqrt(sum(!is.na((!!var)))),
              min = min((!!var),na.rm = T),
              max = max((!!var),na.rm = T),
              IQR = IQR((!!var), na.rm = TRUE),
              Q1 = quantile((!!var), probs = 0.25, na.rm = TRUE),
              Q3 = quantile((!!var), probs = 0.75, na.rm = TRUE),
              "  - Not available" = sum(is.na(!!var)),
              n = n(),
              .groups = "drop") %>% # d <- d_
    round.fun() %>%
    # mean(ci), median(icr) or median(range)
    {if (stat_type == "mean_ci") 
      mutate(.,!!as_label(var) := pmap_chr(., ~mean_confidence.fun(..3, ..4))) else .} %>%
    {if (stat_type == "median_q1q3") 
      mutate(.,!!as_label(var) := paste0(.$median," (",.$Q1,"-",.$Q3,")")) else .} %>%
    {if (grepl("mean_ci|median_range|median_q1q3", stat_type)) 
      mutate(.,!!as_label(var) := paste0(.$median," (",.$min,"-",.$max,")")) else
      stop("Invalid stat_type. Choose 'mean_ci', 'median_q1q3' or 'median_range'.")} %>%
    mutate(percent = round(100*(`  - Not available`/n))) %>%


    mutate("  - Not available" = paste0(.$`  - Not available`," (",.$percent,"%)")) %>%
    mutate_at(vars(`  - Not available`), str_replace_all, pattern = "^0 \\(0%\\)", replacement = "0") %>%
    select(!!group, !!as_label(var), `  - Not available`) %>%
    #dplyr::rename(., !!(var) :="val" ) %>%
    pivot_longer(-!!group, names_to='Variables', values_to='value') %>%
    pivot_wider(names_from = !!group, values_from = value) %>%
    mutate("p-value" = c(p.val, NA)) %>%
    mutate(na = c( sum(is.na(pull(df, !!var))), NA )) %>%
    filter(!if_all(all_of(l), ~grepl("^0$", .) & Variables == "  - Not available")) %>%
    mutate_if(is.character, str_replace_all, pattern = "NA \\(Inf--Inf\\)", replacement = "NA (NA)") %>%
    mutate("t" = t)

  return(sum_tab)
  #return(list(exp1, df, lvl=lvl))
}   

###########################
# COMBINE TABLES FUNCTION #
###########################
# gr_order <- c("NonIners", "Iners", "Gard", "Prev", "Oth")
# d_ <- sum_table.fun(metadata, val, ModifiedM4, gr_order)
# a <- sum_table.fun(meta_all, val, Contraception, gr_order)
sum_table.fun <- function(df, val, group, gr_order) {
  group <- enquo(group)
  
  # round decimals
  round.fun <- function(x){
    df <- x %>%
    mutate_if(
    is.numeric,
    ~ ifelse(.x >= 1|.x==0, round(.x, digits = 0),
            ifelse(.x< 1 & .x >= 0.001, round(.x, digits = 2),
                                          "< 0.001")))
  }
  c <- c()
  f <- c()
  
  for (v in val) {
    if(is.numeric( pull(df, v))) 
      {c<-c(c, v)}
    else{f<-c(f, v)}
  } 
  
  # count (n) in each group
  count <- table( pull(df, !!group) ) 
  count_t <- count %>% paste0("**",names(.),"**","\\\n","n=(", ., ")") %>% set_names(., names(count))  
  names_n <- names(count) %>% set_names(., count_t)
  
  l_c <- imap(c, ~continuous.fun(df, !! rlang::sym(.x), (!!group)) ) %>%
    bind_rows(., )
  l_f <- imap(f, ~factor.fun(df, !! rlang::sym(.x), (!!group)) ) %>%
    bind_rows(.)
  
  table <- bind_rows(l_c, l_f) %>%
    round.fun() %>%
    mutate("p-value" = ifelse(is.na(.$`p-value`),NA, paste0("$$ ",.$`p-value`,"^",.$t," $$"))) %>%
    select("Variables", any_of(gr_order), "p-value", "na", -t) %>%
    mutate_at(vars("na"), ~ as.character(.)) %>%
    dplyr::rename(!!! names_n) %>%
    dplyr::rename("**p-value**"=`p-value`, "**na**"='na')

  #return(list(l_c, l_f, table))
  return(table)
  
}
  

sum_table.fun(df=df2 %>% select(Astrand_rel_VO2, Age, Sex, Period), 
           val = c("Sex", "Age"),
           group = Period,
           gr_order = unique(df2$Period))
```

```{r}
condition <- "candida"
gr_order <- c("yes", "no")
val <- c("age", "Contraception", "HIVstatus", "BV_Diagnosis", "Luminal_gr")


a <- sum_table.fun(meta_all[[2]], val, candida, gr_order)
s <- c(" ", rep(c("median (range or %)"), times=length(gr_order)), NA, NA) %>% set_names(., colnames(a))

a_ <- bind_rows(s, a) %>% 
  setNames(gsub("yes","Candida group",names(.))) %>%
  setNames(gsub("no","Control group",names(.))) %>%
  #add_row(Variables = "Demographics", .before = 1) %>%
  mutate(Variables = str_replace_all(string = .$Variables, pattern = clean_names)) %>%
  mutate(Variables = str_replace_all(string = .$Variables, pattern = bact_names)) %>%
  mutate(Variables = str_replace_all(string = .$Variables, pattern = BV_names)) %>%
  mutate(Variables = ifelse(grepl("  -.+|^ $", .$Variables), 
                            .$Variables, paste0("**",.$Variables,"**")) ) #%>%
  #select(-`**na**`, -`**p-value**`) #%>%
  #column_to_rownames(var = "Variables")

pander(a_, keep.line.breaks = TRUE)
```

```{r}
library(pander)
condition <- "Period"
gr_order <- c("1995-1999", "2000-2004", "2005-2009", "2010-2014", "2015-2019", "2020-2023")  #unique(df2$Period)
val <- c("Sex", "Age", "Astrand_rel_VO2")

a <- df2 %>% 
  select(Astrand_rel_VO2, Age, Sex, Period) %>%
  sum_table.fun( val, Period, gr_order)

pander(a, keep.line.breaks = TRUE)

s <- c(" ", rep(c("median (range or %)"), times=length(gr_order)), NA, NA) %>% set_names(., colnames(a))

a_ <- bind_rows(s, a) %>% 
  mutate(Variables = ifelse(grepl("  -.+|^ $", .$Variables), 
                            .$Variables, paste0("**",.$Variables,"**")) ) #%>%
  #select(-`**na**`, -`**p-value**`) #%>%
  #column_to_rownames(var = "Variables")

pander(a_, keep.line.breaks = TRUE)

DT::datatable(a_)
flextable(a_)
  
```

## -Table 1 Lifestyle charasteristics
 
```{r}
val_use <- 
  colnames(
  df2 %>% select(Age, HeightCM, WeightKG, BMI, Astrand_rel_VO2, EkB_rel_VO2, 
                 BloodPressureSystolic, BloodPressureDiastolic, TobaccoSmoking))
to_char <- c("TobaccoSmoking")

df2 %>% 
  select(Period, any_of(val_use)) %>% 
  mutate(across(any_of(to_char), ~as.character(.))) %>%
  
  sum_table.fun(., 
                val = val_use,
                group = Period,
                gr_order = unique(df2$Period))





# Create summarized table
summary_table <- df2 %>%

  group_by(Period) %>%
  summarise(
    Age = mean(Age, na.rm = TRUE),
    HeightCM = mean(HeightCM, na.rm = TRUE),
    WeightKG = mean(WeightKG, na.rm = TRUE),
    BMI = mean(BMI, na.rm = TRUE),
    BloodPressureSystolicMean = mean(BloodPressureSystolic, na.rm = TRUE),
    BloodPressureDiastolicMean = mean(BloodPressureDiastolic, na.rm = TRUE),
    EkB_rel_VO2 = mean(EkB_rel_VO2, na.rm = TRUE),
    Astrand_rel_VO2 = mean(Astrand_rel_VO2, na.rm = TRUE),
    Smokers = mean(TobaccoSmoking < 2, na.rm = TRUE) * 100
  ) %>% 
  mutate(across(where(is.numeric), round, 2))

summary_table <- df2 %>%
  group_by(Period) %>%
  summarise(
    Count = n(),
    Age_Mean = mean(Age, na.rm = TRUE),
    Age_SD = sd(Age, na.rm = TRUE),
    HeightCM_Mean = mean(HeightCM, na.rm = TRUE),
    HeightCM_SD = sd(HeightCM, na.rm = TRUE),
    WeightKG_Mean = mean(WeightKG, na.rm = TRUE),
    WeightKG_SD = sd(WeightKG, na.rm = TRUE),
    BMI_Mean = mean(BMI, na.rm = TRUE),
    BMI_SD = sd(BMI, na.rm = TRUE),
    BloodPressureSystolicMean = mean(BloodPressureSystolic, na.rm = TRUE),
    BloodPressureSystolic_SD = sd(BloodPressureSystolic, na.rm = TRUE),
    BloodPressureDiastolicMean = mean(BloodPressureDiastolic, na.rm = TRUE),
    BloodPressureDiastolic_SD = sd(BloodPressureDiastolic, na.rm = TRUE),
    EkB_rel_VO2_Mean = mean(EkB_rel_VO2, na.rm = TRUE),
    EkB_rel_VO2_SD = sd(EkB_rel_VO2, na.rm = TRUE),
    Astrand_rel_VO2_Mean = mean(Astrand_rel_VO2, na.rm = TRUE),
    Astrand_rel_VO2_SD = sd(Astrand_rel_VO2, na.rm = TRUE),
    Smokers = mean(TobaccoSmoking < 2, na.rm = TRUE) * 100,
    Smokers_SE = sqrt((Smokers/100) * (1 - Smokers/100) / Count)
  ) %>%
  mutate(
    Age_CI = paste0(round(Age_Mean, 2), " (", round(Age_Mean - 1.96 * Age_SD / sqrt(Count), 2), "-", round(Age_Mean + 1.96 * Age_SD / sqrt(Count), 2), ")"),
    HeightCM_CI = paste0(round(HeightCM_Mean, 2), " (", round(HeightCM_Mean - 1.96 * HeightCM_SD / sqrt(Count), 2), "-", round(HeightCM_Mean + 1.96 * HeightCM_SD / sqrt(Count), 2), ")"),
    WeightKG_CI = paste0(round(WeightKG_Mean, 2), " (", round(WeightKG_Mean - 1.96 * WeightKG_SD / sqrt(Count), 2), "-", round(WeightKG_Mean + 1.96 * WeightKG_SD / sqrt(Count), 2), ")"),
    BMI_CI = paste0(round(BMI_Mean, 2), " (", round(BMI_Mean - 1.96 * BMI_SD / sqrt(Count), 2), "-", round(BMI_Mean + 1.96 * BMI_SD / sqrt(Count), 2), ")"),
    BloodPressureSystolic_CI = paste0(round(BloodPressureSystolicMean, 2), " (", round(BloodPressureSystolicMean - 1.96 * BloodPressureSystolic_SD / sqrt(Count), 2), "-", round(BloodPressureSystolicMean + 1.96 * BloodPressureSystolic_SD / sqrt(Count), 2), ")"),
    BloodPressureDiastolic_CI = paste0(round(BloodPressureDiastolicMean, 2), " (", round(BloodPressureDiastolicMean - 1.96 * BloodPressureDiastolic_SD / sqrt(Count), 2), "-", round(BloodPressureDiastolicMean + 1.96 * BloodPressureDiastolic_SD / sqrt(Count), 2), ")"),
    EkB_rel_VO2_CI = paste0(round(EkB_rel_VO2_Mean, 2), " (", round(EkB_rel_VO2_Mean - 1.96 * EkB_rel_VO2_SD / sqrt(Count), 2), "-", round(EkB_rel_VO2_Mean + 1.96 * EkB_rel_VO2_SD / sqrt(Count), 2), ")"),
    Astrand_rel_VO2_CI = paste0(round(Astrand_rel_VO2_Mean, 2), " (", round(Astrand_rel_VO2_Mean - 1.96 * Astrand_rel_VO2_SD / sqrt(Count), 2), "-", round(Astrand_rel_VO2_Mean + 1.96 * Astrand_rel_VO2_SD / sqrt(Count), 2), ")"),
    Smokers_CI = paste0(round(Smokers, 2), " (", round(Smokers - 1.96 * Smokers_SE, 2), "-", round(Smokers + 1.96 * Smokers_SE, 2), ")")
  ) %>%
  select(Period, Count, Age_CI, HeightCM_CI, WeightKG_CI, BMI_CI, BloodPressureSystolic_CI, BloodPressureDiastolic_CI, EkB_rel_VO2_CI, Astrand_rel_VO2_CI, Smokers_CI,
         -Count) 

 summary_table_transposed <- 
   summary_table %>%
  pivot_longer(
    -Period, 
    names_to = "Attribute", 
    values_to = "Value"
  ) %>%
  pivot_wider(
    names_from = Period, 
    values_from = Value,
    names_prefix = "Period_"
   ) 
  
 # %>% 
  #  mutate(across(where(is.numeric), round,1)) %>% 
  #  flextable::flextable() 


# Skapa en DataTable från den sammanfogade datan
datatable(
  summary_table_transposed,
  extensions = 'Buttons',
  options = list(
    scrollX = TRUE,
    pageLength = 25,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    title = "Sammanställd Frekvens av ICD-koder"
  )
)



```

## -Table 2  - Sociodemographics
N och % i var grupp


Income level (% of median) 
≥200 
120–199 
80–119 
60–79 
<60 


  
```{r}

val_use <- 
  colnames(
  df2 %>% select(
    Sex, AgeGroup, EducationLevel, KommunSize, MarriedOrPartner, BirthPlace
    )
  )

t2 <- df2 %>% 
  select(Period, any_of(val_use)) %>% 
  mutate(across(any_of(to_char), ~as.character(.))) %>%
    mutate(
    EducationLevel = 
      fct_relevel(EducationLevel, 
                  "Primary", "Secondary", "Tertiary <2 years", "Higher Education", "Unknown"),
    KommunSize = 
      fct_relevel(KommunSize, 
                  "rural_municipalities", "dense_municipalities", "metropolitan_municipalities", "Other"),
    MarriedOrPartner = 
      fct_relevel(MarriedOrPartner, 
                  "0", "1"),
    BirthPlace = 
      fct_relevel(BirthPlace, 
                  "Sweden", "In Europe", "Outside Europe")
  ) %>% 
  
  sum_table.fun(., 
                val = val_use,
                group = Period,
                gr_order =  c("1995-1999", "2000-2004", "2005-2009", "2010-2014", "2015-2019", "2020-2023")) #unique(df2$Period))




 summary_table <- df2 %>%
  group_by(Period) %>%
  summarise(
    n = formatC(n(), big.mark = ","),
    Men = sprintf("%s (%.1f%%)", formatC(sum(Sex == "Men"), big.mark = ","), mean(Sex == "Men") * 100),
    Age18_35 = sprintf("%s (%.1f%%)", formatC(sum(AgeGroup == "18-35"), big.mark = ","), mean(AgeGroup == "18-35") * 100),
    Age36_50 = sprintf("%s (%.1f%%)", formatC(sum(AgeGroup == "36-50"), big.mark = ","), mean(AgeGroup == "36-50") * 100),
    Age51_65 = sprintf("%s (%.1f%%)", formatC(sum(AgeGroup == "51-65"), big.mark = ","), mean(AgeGroup == "51-65") * 100),
    AgeOLD = sprintf("%s (%.1f%%)", formatC(sum(AgeGroup == ">65"), big.mark = ","), mean(AgeGroup == ">65") * 100),
    MarriedOrPartner = sprintf("%s (%.1f%%)", formatC(sum(MarriedOrPartner == 1), big.mark = ","), mean(MarriedOrPartner == 1) * 100),
    BirthPlaceSweden = sprintf("%s (%.1f%%)", formatC(sum(BirthPlace == "Sweden"), big.mark = ","), mean(BirthPlace == "Sweden") * 100),
    BirthPlaceInEurope = sprintf("%s (%.1f%%)", formatC(sum(BirthPlace == "In Europe"), big.mark = ","), mean(BirthPlace == "In Europe") * 100),
    BirthPlaceOutsideEurope = sprintf("%s (%.1f%%)", formatC(sum(BirthPlace == "Outside Europe"), big.mark = ","), mean(BirthPlace == "Outside Europe") * 100),
    White_collar = sprintf("%s (%.1f%%)", formatC(sum(SSYK_WB == "White-Collar"), big.mark = ","), mean(SSYK_WB == "White-Collar") * 100),
    Blue_collar = sprintf("%s (%.1f%%)", formatC(sum(SSYK_WB == "Blue-Collar"), big.mark = ","), mean(SSYK_WB == "Blue-Collar") * 100),
    High_Skilled = sprintf("%s (%.1f%%)", formatC(sum(SSYK_Category == "High-Skilled", na.rm = TRUE), big.mark = ","), mean(SSYK_Category == "High-Skilled", na.rm = TRUE) * 100),
    Clerical_and_Service = sprintf("%s (%.1f%%)", formatC(sum(SSYK_Category == "Clerical and Service Workers", na.rm = TRUE), big.mark = ","), mean(SSYK_Category == "Clerical and Service Workers", na.rm = TRUE) * 100),
    Skilled_Manual = sprintf("%s (%.1f%%)", formatC(sum(SSYK_Category == "Skilled Manual", na.rm = TRUE), big.mark = ","), mean(SSYK_Category == "Skilled Manual", na.rm = TRUE) * 100),
    Elementary_Occupations = sprintf("%s (%.1f%%)", formatC(sum(SSYK_Category == "Elementary Occupations", na.rm = TRUE), big.mark = ","), mean(SSYK_Category == "Elementary Occupations", na.rm = TRUE) * 100),
    metropolitan_municipalities = sprintf("%s (%.1f%%)", formatC(sum(KommunSize == "metropolitan_municipalities"), big.mark = ","), mean(KommunSize == "metropolitan_municipalities") * 100),
    dense_municipalities = sprintf("%s (%.1f%%)", formatC(sum(KommunSize == "dense_municipalities"), big.mark = ","), mean(KommunSize == "dense_municipalities") * 100),
    rural_municipalities = sprintf("%s (%.1f%%)", formatC(sum(KommunSize == "rural_municipalities"), big.mark = ","), mean(KommunSize == "rural_municipalities") * 100),
    EducationPrimary = sprintf("%s (%.1f%%)", formatC(sum(EducationLevel == "Primary"), big.mark = ","), mean(EducationLevel == "Primary") * 100),
    EducationSecondary = sprintf("%s (%.1f%%)", formatC(sum(EducationLevel == "Secondary"), big.mark = ","), mean(EducationLevel == "Secondary") * 100),
    EducationTertiary = sprintf("%s (%.1f%%)", formatC(sum(EducationLevel == "Tertiary <2 years"), big.mark = ","), mean(EducationLevel == "Tertiary <2 years") * 100),
    EducationHigher = sprintf("%s (%.1f%%)", formatC(sum(EducationLevel == "Higher Education"), big.mark = ","), mean(EducationLevel == "Higher Education") * 100),
    IncomeLess60 = sprintf("%s (%.1f%%)", formatC(sum(IncomeLevel == "<60%"), big.mark = ","), mean(IncomeLevel == "<60%") * 100),
    Income60_79 = sprintf("%s (%.1f%%)", formatC(sum(IncomeLevel == "60–79%"), big.mark = ","), mean(IncomeLevel == "60–79%") * 100),
    Income80_119 = sprintf("%s (%.1f%%)", formatC(sum(IncomeLevel == "80–119%"), big.mark = ","), mean(IncomeLevel == "80–119%") * 100),
    Income120_199 = sprintf("%s (%.1f%%)", formatC(sum(IncomeLevel == "120–199%"), big.mark = ","), mean(IncomeLevel == "120–199%") * 100),
    Income200Plus = sprintf("%s (%.1f%%)", formatC(sum(IncomeLevel == "≥200%"), big.mark = ","), mean(IncomeLevel == "≥200%") * 100)
  ) %>%
  pivot_longer(
    cols = -Period, 
    names_to = "Attribute", 
    values_to = "Value"
  ) %>%
  pivot_wider(
    names_from = Period, 
    values_from = Value,
    names_prefix = "Period_"
  ) %>% 
  mutate(
    Table_Category = case_when(
      Attribute %in% c("Men", "Age18_35", "Age36_50", "Age51_65", "AgeOLD", "Married", "BirthPlaceSweden", "BirthPlaceInEurope", "BirthPlaceOutsideEurope", "EducationPrimary", "EducationSecondary", "EducationTertiary", "EducationHigher") ~ "Sociodemographic",
      TRUE ~ "Worklife Characteristics"
    )
  ) %>%
  flextable::flextable()
```



# finction
Handles both continuous (numeric) and categorical data: For continuous data, it should calculate either mean with a 95% confidence interval or median with interquartile range (IQR), based on the specified method. For categorical data, it should count each category and calculate the percentage.

Groups by a specified column and should be able to use without groupby: The function should group the data by this column and perform the summarizations within each group.

Outputs a wide-format table: Each group should be a separate column, and each row should correspond to a specific variable or category.

Modular approach: Separate helper functions for numeric and categorical data, allowing for different summary types.
```{r}
create_summary_table <- function(df, numeric_summary = c("mean", "ci"), categorical_summary = "n_pct", group_by_column = NULL) {
  
  # Helper function for numeric summary
  numeric_summary_fun <- function(x, summary_type) {
    if (summary_type == "mean") {
      mean(x, na.rm = TRUE)
    } else if (summary_type == "median") {
      median(x, na.rm = TRUE)
    } else if (summary_type == "ci") {
      # Calculate mean and 95% confidence interval
      m <- mean(x, na.rm = TRUE)
      se <- sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))
      ci_lower <- m - 1.96 * se
      ci_upper <- m + 1.96 * se
      paste0(round(m, 2), " (", round(ci_lower, 2), "-", round(ci_upper, 2), ")")
    } else {
      NA
    }
  }

  # Helper function for categorical summary
  categorical_summary_fun <- function(x) {
    if (categorical_summary == "n_pct") {
      sprintf("%s (%.1f%%)", formatC(sum(x), big.mark = ","), 100 * mean(x, na.rm = TRUE))
    } else {
      NA
    }
  }
  
  # Conditional grouping
  if (!is.null(group_by_column)) {
    df <- df %>% group_by(!!sym(group_by_column))
  }
  
  # Create summarized table
  summary_table <- df %>%
    summarise(across(where(is.numeric), ~numeric_summary_fun(., numeric_summary)),
              across(where(is.factor), ~categorical_summary_fun(.)))
  
  # Transpose if needed
  summary_table_transposed <- summary_table %>%
    pivot_longer(cols = -!!sym(group_by_column), names_to = "Attribute", values_to = "Value") %>%
    pivot_wider(names_from = !!sym(group_by_column), values_from = Value, names_prefix = paste0(group_by_column, "_"))

  return(summary_table_transposed)
}

# Example usage
 result <- create_summary_table(df2 %>% select(EkB_rel_VO2,Age, Sex, Period), numeric_summary = "ci", group_by_column = "Period")
 
 
 
 
 
 
 
 

 
 
library(dplyr)
library(tidyr)

create_summary_table <- function(df, numeric_summary = c("mean", "ci"), categorical_summary = "n_pct", group_by_column = NULL) {
  
  # Helper function for numeric summary
  numeric_summary_fun <- function(x, summary_types) {
    summary_values <- character()
    if ("mean" %in% summary_types) {
      summary_values <- c(summary_values, round(mean(x, na.rm = TRUE), 2))
    }
    if ("median" %in% summary_types) {
      summary_values <- c(summary_values, round(median(x, na.rm = TRUE), 2))
    }
    if ("ci" %in% summary_types) {
      m <- mean(x, na.rm = TRUE)
      se <- sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))
      ci_lower <- round(m - 1.96 * se, 2)
      ci_upper <- round(m + 1.96 * se, 2)
      summary_values <- c(summary_values, paste0(m, " (", ci_lower, "-", ci_upper, ")"))
    }
    return(toString(summary_values))
  }

  # Helper function for categorical summary
  categorical_summary_fun <- function(x) {
    if (categorical_summary == "n_pct") {
      tbl <- table(x)
      pct <- prop.table(tbl) * 100
      return(toString(paste(names(tbl), formatC(tbl, big.mark = ","), "(", round(pct, 1), "%)", sep=": ")))
    }
    return(NA)
  }
  
  # Conditional grouping
  if (!is.null(group_by_column)) {
    df <- df %>% group_by(!!sym(group_by_column))
  }
  
  # Create summarized table
  summary_table <- df %>%
    summarise(
      across(where(is.numeric), ~numeric_summary_fun(., numeric_summary), .names = "num_{.col}"),
      across(where(~ is.factor(.x) || is.character(.x)), ~categorical_summary_fun(.x), .names = "cat_{.col}")
    )

  # Pivot and format the summary table
  summary_table_transposed <- summary_table %>%
    pivot_longer(cols = -!!sym(group_by_column), names_to = "Attribute", values_to = "Value") %>%
    pivot_wider(names_from = Attribute, values_from = Value, names_prefix = "")

  return(summary_table_transposed)
}

# Example usage
result <- create_summary_table(df2 %>% select(EkB_rel_VO2, Age, Sex, Period), group_by_column = "Period")

# Example usage
 result <- create_summary_table(df2 %>% select(EkB_rel_VO2,Age, Sex, Period), numeric_summary = "mean_ci", group_by_column = "Period", round_digits = 2)

# Example usage
 result <- create_summary_table(df2 %>% select(EkB_rel_VO2,Age, Sex, Period), numeric_summary = "mean_ci", group_by_column = "Period")

# Example usage
 result_without_group <- create_summary_table(df2 %>% select(EkB_rel_VO2,Age, Sex, Period), numeric_summary = "median_iqr")
 result_with_group <- create_summary_table(df2 %>% select(EkB_rel_VO2,Age, Sex, Period), numeric_summary = "mean_ci", group_by_column = "Period")
# Example usage
 result <- create_summary_table(df2 %>% select(EkB_rel_VO2,Age, Sex, Period), numeric_summary = "median_iqr",categorical_summary = "n_pct", group_by_column = "Period")
 
 
 
 
 
 
 
 
library(tidyverse)

create_summary_table <- function(df, numeric_cols, cat_cols, group_by_col=NULL, numeric_summary="mean_ci") {
  
  # Helper function for numeric summary (mean and 95% CI)
  numeric_summary_fun <- function(x, summary_type) {
    if (summary_type == "mean_ci") {
      m <- mean(x, na.rm = TRUE)
      se <- sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))
      ci_lower <- round(m - 1.96 * se, 2)
      ci_upper <- round(m + 1.96 * se, 2)
      return(paste0(m, " (", ci_lower, "-", ci_upper, ")"))
    }
    return(NA)
  }

  # Helper function for categorical summary (count and percentage)
  categorical_summary_fun <- function(x) {
    tbl <- table(x)
    pct <- prop.table(tbl) * 100
    return(data.frame(category = names(tbl), count = tbl, percent = round(pct, 1)))
  }
  
  # Conditional grouping
  if (!is.null(group_by_col)) {
    df <- df %>% group_by(!!sym(group_by_col))
  }
  
  # Create summarized table
  numeric_summary_df <- df %>%
    summarise(across(all_of(numeric_cols), ~numeric_summary_fun(., numeric_summary), .names = "num_{.col}"))

  cat_summary_df <- map_dfr(all_of(cat_cols), ~{
    df %>%
      summarise(!!sym(.x) := categorical_summary_fun(!!sym(.x))) %>%
      unnest(!!sym(.x))
  }, .id = "category")

  # Combine numeric and categorical summaries
  combined_summary_df <- bind_rows(numeric_summary_df, cat_summary_df)

  # Pivot and format the summary table
  summary_table_transposed <- combined_summary_df %>%
    pivot_longer(cols = -!!sym(group_by_col), names_to = "Attribute", values_to = "Value") %>%
    pivot_wider(names_from = !!sym(group_by_col), values_from = Value)

  return(summary_table_transposed)
}

# Example usage:
# create_summary_table(df2, numeric_cols = c("Age", "Astrand_rel_VO2"), cat_cols = c("Sex"), group_by_col = "Period", numeric_summary = "mean_ci")

# Example usage:
create_summary_table(df2 %>% select(Astrand_rel_VO2, Age, Sex, Period), numeric_cols = c("Age", "Astrand_rel_VO2"), cat_cols = c("Sex"), group_by_col = "Period", numeric_summary = "mean_ci")



# Example usage:
 create_summary_table(df2 %>% select(Astrand_rel_VO2, Age, Sex, Period ), numeric_cols = c("Age", "Astrand_rel_VO2"), cat_cols = c("Sex"), group_by_col = "Period", numeric_summary = "mean_ci")
# create_summary_table(df, numeric_cols = c("age", "height"), cat_cols = c("gender", "smoker"), numeric_summary = "median_iqr")

```







## -Table 3 worklife charasteristics


Total
Contractual relationship insecurity
Directly employed by the employer (0)
Employed by an agency (-1) Combination of self- and direct employment (-1) Self-employed (-1)
Solo self-employed (-2)
Contractual temporariness Stable employment (0) Unstable employment (-2)
Multiple jobs/economic sectors
CBA coverage (% likelihood)
https://isf.se/publikationer/rapporter/2018/2018-09-04-vem-far-avsattningar-till-tjanstepension
```{r}

# Create summarized table
summary_table <- df2 %>%

  group_by(Period) %>%
  summarise(
    n = n(),
    Men = mean(Sex == "Men") * 100,
    Age18_35 = mean(AgeGroup == "18-35") * 100,
    Age36_50 = mean(AgeGroup == "36-50") * 100,
    Age51_65 = mean(AgeGroup == "51-65") * 100,
    AgeOLD = mean(AgeGroup == ">65") * 100,
    Married = mean(Married, na.rm = TRUE) * 100,
    BirthPlaceSweden = mean(BirthPlace == "Sweden") * 100,
    BirthPlaceInEurope = mean(BirthPlace == "In Europe") * 100,
    BirthPlaceOutsideEurope = mean(BirthPlace == "Outside Europe") * 100,
    White_collar = mean(SSYK_WB == "White-Collar") * 100,
    Blue_collar = mean(SSYK_WB == "Blue-Collar") * 100,
    metropolitan_municipalities = mean(KommunSize == "metropolitan_municipalities") * 100,
    dense_municipalities = mean(KommunSize == "dense_municipalities") * 100,
    rural_municipalities = mean(KommunSize == "rural_municipalities") * 100,
    # MedianIncome = median(CSFVI),
    MedianIncomeQ1 = median(CSFVI[IncomeQuartile == 1], na.rm = TRUE),
    MedianIncomeQ2 = median(CSFVI[IncomeQuartile == 2], na.rm = TRUE),
    MedianIncomeQ3 = median(CSFVI[IncomeQuartile == 3], na.rm = TRUE),
    MedianIncomeQ4 = median(CSFVI[IncomeQuartile == 4], na.rm = TRUE),
    EducationLow = mean(EducationLevel == "Basic education") * 100,
    EducationMedium = mean(EducationLevel == "Secondary to Post-Secondary Education") * 100,
    EducationHigh = mean(EducationLevel == "Higher Education") * 100,
    BloodPressureSystolicMean = mean(BloodPressureSystolic, na.rm = TRUE),
    BloodPressureDiastolicMean = mean(BloodPressureDiastolic, na.rm = TRUE),
    BMI = mean(BMI, na.rm = TRUE),
    EkB_rel_VO2 = mean(EkB_rel_VO2, na.rm = TRUE),
    Astrand_rel_VO2 = mean(Astrand_rel_VO2, na.rm = TRUE),
    Smokers = mean(TobaccoSmoking < 2) * 100
  ) %>% 
  mutate(across(where(is.numeric), round, 2))



 summary_table_transposed <- 
   summary_table %>%
  pivot_longer(
    -Period, 
    names_to = "Attribute", 
    values_to = "Value"
  ) %>%
  pivot_wider(
    names_from = Period, 
    values_from = Value,
    names_prefix = "Period_"
   ) 
  
 # %>% 
  #  mutate(across(where(is.numeric), round,1)) %>% 
  #  flextable::flextable() 


# Skapa en DataTable från den sammanfogade datan
datatable(
  summary_table_transposed,
  extensions = 'Buttons',
  options = list(
    scrollX = TRUE,
    pageLength = 25,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    title = "Sammanställd Frekvens av ICD-koder"
  )
)

```




#-------------------------
# HPI vs SCB data

## -Table xx

```{r}

```


## -Figure xx


#-------------------------
# 30 most underrepresented/overrepresented groups

```{r}

```


## -Figure xx
```{r}

```

#-------------------------
# other stuff not included in paper

## Income graph
```{r}

df1 %>%
  mutate(Period = case_when(
    Year >= 1995 & Year < 2000 ~ "1995-1999",
    Year >= 2000 & Year < 2005 ~ "2000-2004",
    Year >= 2005 & Year < 2010 ~ "2005-2009",
    Year >= 2010 & Year < 2015 ~ "2010-2014",
    Year >= 2015 & Year < 2020 ~ "2015-2019",
    Year >= 2020 & Year <= 2023 ~ "2020-2023",
    TRUE ~ NA_character_  # For years outside the specified range
  )) %>%
  ggplot(aes(x = Period, y = CSFVI)) + 
  geom_violin() +
  geom_boxplot(width = 0.2, 
                 outlier.color = "red",
  outlier.fill = "red",
  outlier.shape = 19,
  outlier.size = .3,
  outlier.stroke = 0.5,
  outlier.alpha = .8,
  position = position_dodge() )+  # Still hide outliers in boxplot for clarity
  #geom_jitter(color = "red", size = 1, width = 0.1, height = 0, aes(group = Period)) +  # Add jitter for outliers
  scale_y_log10(labels = scales::label_number(), breaks = scales::log_breaks(n = 20)) +  # More breaks on the y-axis
  theme_minimal() +
  labs(y = "CSFVI (log scale)", x = "Period")

```
